---
title: "Jet Blue reach to the top"
author: "Nishan Madawanarachchi"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data
The data provided are the domestic market data reported for US carriers by Bureau of Transportation Statistics; availble [*here*](https://www.transtats.bts.gov/DatabaseInfo.asp?DB_ID=111). We got 28 csv files representing market data from 1990 to 2017.

```{r, warning=F, message=F}
# loading the required librarys
library(magrittr)
library(dplyr)
library(ggplot2)
library(knitr)
```

```{r, cache=T}
#Getting the CSV files except 2017 data as it is not completed
flights <- data.frame(NULL)
for(i in 1990:2016) {  
    file <- paste0("657240010_T_T100D_MARKET_US_CARRIER_ONLY_",i,"_All.csv")
    flights <- rbind(flights, read.csv(paste0("q3_data/",file)) )
}
```

```{r}
str(flights)
```

Dataframe `flights` has `r nrow(flights)` rows and `r ncol(flights)` variables.

## Cleaning up data for unwanted columns
```{r, cache=T}
cols <- c("YEAR", "MONTH", "PASSENGERS", "FREIGHT", "MAIL", "CARRIER_GROUP_NEW", "UNIQUE_CARRIER_NAME")
flights2 <- flights[cols]
kable(head(flights2))
```

## Lets first see how enplanned number of passengers grew over the years

```{r,cache=T}
# making year a factor
flights2$YEAR <- as.factor(flights2$YEAR)

# summing enplaned passengers, freight and mail per year basis
byyear_pfm <- flights2 %>% group_by(YEAR) %>% 
    summarise(PASSENGERS = sum(PASSENGERS,na.rm = T), FREIGHT = sum(FREIGHT, na.rm = T), 
              MAIL = sum(MAIL, na.rm = T))
kable(head(flights2))
```

```{r, cache=T, fig.height=2, fig.width= 4}
# plot enplanned passenger growth
ggplot(byyear_pfm, aes(x = YEAR, y = PASSENGERS, group = 1)) + geom_line() + 
    theme(text = element_text(size=10),axis.text.x = element_text(angle=90))
```

Enplaned number of passengers has steadily grown over the years from 1990 to 2016.
However, there are decrease in the number of passengers in the period 2000-2002 as 
well as in the period 2007-2009.

## Lets see the mix of the carrier groups interms of number of passengers in 2016

```{r, cache=T}
# making CARRIER_GROUP_NEW a factor variable
flights2$CARRIER_GROUP_NEW <- as.factor(flights2$CARRIER_GROUP_NEW)
# looking at 2016 carrier group mix
carriergrp_2016 <- flights2 %>% filter(YEAR == 2016, PASSENGERS > 0) %>% 
    group_by(CARRIER_GROUP_NEW) %>% summarise(PASSENGERS = sum(PASSENGERS,na.rm = T))
```

```{r, cache=T}
carriergrp_2016 %<>% mutate(Share = round(PASSENGERS*100/sum(PASSENGERS), 2) )
kable(carriergrp_2016)
```

From the table above it can be seen that Major Carriers (3) has highest share (88.16%) 
of enplanned passengers. Well behind major carriers are the national carriers (2) with a
share of (10.14%)

## Lets see how the percentage share between Major and national carriers changed over the years

```{r, cache = T}
carriergrp <- flights2 %>% filter(CARRIER_GROUP_NEW == 2 | CARRIER_GROUP_NEW == 3, 
                                       PASSENGERS > 0) %>% 
    group_by(YEAR, CARRIER_GROUP_NEW) %>% summarise(PASSENGERS = sum(PASSENGERS,na.rm = T))
```

```{r, cache=T, fig.height=3, fig.width= 6}
# calculating share between national carriers and major carriers
carriergrp %<>% mutate(share = round(PASSENGERS*100/sum(PASSENGERS), 2) )
ggplot(carriergrp, aes(x = YEAR, y = share, col = CARRIER_GROUP_NEW, group = 1)) + 
    geom_point( ) + theme(text = element_text(size=10),axis.text.x = element_text(angle=90))

# maximum passenger share of national carriers
maxshare <- max(filter(carriergrp, CARRIER_GROUP_NEW==2)$share)
```

It can be the seen from the graph above that national carriers were making ground
interms of passenger share in the period 1990 to 2004, where it reach the maximum
passenger share of `r maxshare` compared to major carriers. Afterwards the share 
of national carriers hits a downward trend. However, it can be assumed that during 
the period 2004-2008 some of the national carriers were perfoming well interms of 
revenue and reclassifed as major carriers.

## Lets check for years where national carriers were reclassifed as major carriers

```{r, cache = T}
# data frame of national and major carriers
natMajflights <- flights2 %>% select(YEAR, MONTH, UNIQUE_CARRIER_NAME, 
                                     CARRIER_GROUP_NEW) %>%
    filter(CARRIER_GROUP_NEW == 2 | CARRIER_GROUP_NEW == 3, MONTH == 1)
```

```{r, cache=T}
# table with unqiue group and carrier values
natMajflights2 <- natMajflights %>% select(UNIQUE_CARRIER_NAME, CARRIER_GROUP_NEW) %>%
    distinct()
```

```{r, chache=T}
# changed list carriers
changedlist <- natMajflights2[duplicated(natMajflights2$UNIQUE_CARRIER_NAME), ]
# we are only interested in 2 to 3
changedlist <- as.character(changedlist[changedlist$CARRIER_GROUP_NEW == 3,1])
```

```{r, chache=T}
#reducing natMajflights to only group 3 carriers
natMajflights <- dplyr::filter(natMajflights, CARRIER_GROUP_NEW == 3)
indices <- match(changedlist, natMajflights$UNIQUE_CARRIER_NAME)
```

```{r}
# Table of years when recalssification happened
kable(data.frame(Carrier = changedlist, year = natMajflights$YEAR[indices] ))
```

It is clearly visible from the above table that 9 out of the national carriers 
were reclassified as Major carriers during the period 2004-2008 indicating strong 
growth in these carriers.

## Lets see the most popular Major carriers among the passengers in last 5 years

I am assuming the most popular carriers are the ones with highest enplaned passengers
```{r, cache=T}
# sum of passengers by carrier
top5 <- flights2 %>% dplyr::filter(YEAR == 2012 | YEAR == 2013 |YEAR == 2014 |
                                       YEAR == 2015 | YEAR == 2016) %>% 
    group_by(YEAR, UNIQUE_CARRIER_NAME) %>% summarise(PASSENGERS = sum(PASSENGERS)) %>%
    ungroup %>% arrange(YEAR, desc(PASSENGERS))
```

```{r, cache = T}
top5_ref <- list()
j=1
for(i in 2012:2016) {
    temp <- dplyr::filter(top5, YEAR == i)[1:5,2]
    top5_ref[[j]] <- temp
    j <- j + 1
}
```

```{r, cache = T}
top5_ref <- as.data.frame(top5_ref)
colnames(top5_ref) <- c("2012","2013","2014","2015","2016")
kable(top5_ref[,1:3])
kable(top5_ref[,4:5])
```

It can be seen from the tables above that jet blue airways had made giant strides in
securing a top 5 spot in 2015 and 2016 (since becoming classified as major airline in 
2005) surpassing US Airways Inc. 