---
title: "Does race affect the result of traffic stops"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data

This dataset contains yearly traffic stops information provided by Charlotte-Mecklenburg Police Department(CMPD)  available [*here*](https://catalog.data.gov/dataset/officer-traffic-stops).


```{r, warning=F, message=F}
# Loading libraries needed
library(tidyverse)
library(knitr)
library(magrittr)
```

Loading the the `Officer_Traffic_Stops.csv` data
```{r, cache=T}
trfic_stps <- read.csv("Officer_Traffic_Stops.csv")

# taking unwanted columns out
trfic_stps <- trfic_stps[, !names(trfic_stps) %in% 
                             c("ObjectID", "CreationDate", "Creator", "EditDate",
                               "Editor")]
```

```{r}
str(trfic_stps)
```

Dataframe `trfic_stps` has `r nrow(trfic_stps)` rows and `r ncol(trfic_stps)` variables.

## Lets see the reason for stopping vehicles by CMPD in year 2016
```{r, cache=T}
# count of reasons
stops <- as.data.frame(table(trfic_stps$Reason_for_Stop))
top3rsns <- as.character(arrange(stops, desc(Freq))$Var1[1:3])
```


```{r, cache=T}
# bar plot of reasons
stops %>% ggplot(aes(x = reorder(Var1, desc(Freq)), y = Freq)) +
    geom_bar(stat = "Identity") + 
    theme(axis.text.x = element_text(angle=45, size = 8)) + xlab("Reason")
```

It can be seen from the above plot that main 3 causes for vehicle stoppping by CPMD is
due to violations in vehicle registration, speeding and Stop light/sign. 

## Lets understand actions taken by CMPD for each of reason of stopping.

We will focus on top 3 actions taken by CMPD for each stopping event
```{r, cache=T}
# table with reason and result
consistncy <- as.data.frame(table(trfic_stps$Reason_for_Stop, 
                                  trfic_stps$Result_of_Stop))
# taking only first 3 results
consistncy %<>% group_by(Var1) %>% mutate(prop = Freq/sum(Freq)) %>% 
    arrange(Var1, desc(Freq)) %>% slice(1:3)
```

```{r}
# top 3 actions
top3rsn_rslt <- c(0,0,0)
for(i in 1:3)
    top3rsn_rslt[i] <- filter(consistncy, Var1 == top3rsns[i])[1,2]
top3rsn_rslt <- as.character(unlist(top3rsn_rslt))
```

```{r, cache=T}
ggplot(consistncy, aes(x = Var1, y = prop, fill = Var2)) + 
    geom_bar(stat = "Identity") + 
    theme(text = element_text(size=10),axis.text.x = element_text(angle=90)) +
    ylab("Propotion") + xlab("Reason") + 
    scale_fill_discrete(name = "Action taken") + 
    theme(axis.text.x = element_text(angle=45, size = 8)) 
```

Looking at the graph above it seems that there are quite bit variation in the action taken by CPMD. Still it can argued that action taken depends on the situation. Therefore it is worth looking at the profile of the driver.

However, before moving forward I am going to assume that CMPD take following actions  
regarding top 3 traffic violations in general. This will be the baseline.

```{r, cache=T}
kable(data.frame(Reson = top3rsns, General_result = top3rsn_rslt),caption = 'General action taken for top 3 stopping events')
```

## Lets see whether driver profile (Race, Gender, Age) had any impact on the action taken by CPMD

```{r, cache=T}
# filter out the data for only top3 reasons
# create variable to identify consistent (0), and inconsisten (1) actions

new_trfic_stps <- trfic_stps %>% filter(Reason_for_Stop %in% top3rsns) %>%
    mutate(consistent = ifelse( (
        (Reason_for_Stop == top3rsns[1] & Result_of_Stop == top3rsn_rslt[1])|
        (Reason_for_Stop == top3rsns[2] & Result_of_Stop == top3rsn_rslt[2])|
        (Reason_for_Stop == top3rsns[3] & Result_of_Stop == top3rsn_rslt[3])
        ), yes = 0, no = 1))
```

```{r, cache=T}
# general rate being consistent
consisrate <- mean(new_trfic_stps$consistent == 0)

# for gender
consisrate_gender <- new_trfic_stps %>% group_by(Driver_Gender) %>% 
    summarise(rate = mean(consistent == 0))

tdf1 <- data.frame(Variable = rep("Gender", nrow(consisrate_gender)), 
                   Value = unlist(consisrate_gender[,2]), 
                          Profile = unlist(consisrate_gender[,1]))

# for Driver_Race
consisrate_race <- new_trfic_stps %>% group_by(Driver_Race) %>% 
    summarise(rate = mean(consistent == 0))

tdf2 <- data.frame(Variable = rep("Race", nrow(consisrate_race)), 
                   Value = unlist(consisrate_race[,2]), 
                          Profile = unlist(consisrate_race[,1]))

# creating gender groups
new_trfic_stps <- cbind(new_trfic_stps, Age_group = 
                        cut(new_trfic_stps$Driver_Age, 
                            breaks =  c(0, 18, 55, 200), 
                            labels = c("Junior", "Mid Age", "Senior"), 
                            right = F))

consisrate_age <- new_trfic_stps %>% group_by(Age_group) %>% 
    summarise(rate = mean(consistent == 0))

tdf3 <- data.frame(Variable = rep("Age_group", nrow(consisrate_age)), 
                   Value = unlist(consisrate_age[,2]), 
                          Profile = unlist(consisrate_age[,1]) )

# driver profiles
drivr_prof <- rbind(tdf1, tdf2, tdf3)
```

On top 3 traffic violations `r round(consisrate,2)*100`% of the time, action taken by CMPD is consistent with the general actions pointed out in `Table 1`. Now we are going to see how this consistentancy is affected with different driver profiles. So for e.g. if the calculated value is greater than `r round(consisrate,2)`, this means CMPD is more likely to take the general action for this particular driver group. **This will be identified as a consistent behavior**. And if the calculated value is less than `r round(consisrate,2)`, this means CMPD is compartively less likely to take the general action for this particular driver group. **This will be identified as a inconsistent behavior**

```{r, cache=T}
# plot by driver profile
ggplot(drivr_prof, aes(x = Variable, y = Value, group = 1, col = Profile)) + 
    geom_point() + geom_hline(aes(yintercept = consisrate, 
                                  col = "overall consistency rate"), 
                              size=0.25, type = "b") 
    
```

It is visible from the above plot that actions taken by CMPD is more or less consistent for the gender of a driver. Relative to overall consistency rate, it seems that CMPD has been less 
consistent on races of white and native americans, where as they are more consistent
on Asian and Black races. Also, relative to overall consistency rate, it seems that CMPD is more consistent on juniors where as they are less consistent on seniors. It will be interesting to know where these inconsistancies are leaning towards, i.e. whether CPMD being strict or lenient towrds these groups.    

On the other hand tt will be interesting look at profile of police officers to see whether it had a
part to play in CMPD being less or more consistent.

## Lets see whether how actions taken varies with the race of the driver and the profile of the police officer

Profile of the police offcier is provided using attibutes such as race, gender, years of service of the police officer.

```{r, cache=T}
# Race distibution of the driver
kable(count(x = new_trfic_stps, Driver_Race))
```

Table above shows the race distribution of the drivers. White and Black are the larger 
majority. Hence we will focus whether reltively higher consistency towards race of black we saw previously
has anything to do with the profile of the police officer.

```{r, cache=T}
# filter only for black
new_trfic_stps_B <- filter(new_trfic_stps, Driver_Race == "Black")
```

```{r,cache=T}
# overall consistency rate for black
consisrate_B <- mean(new_trfic_stps_B$consistent == 0)

# by gender
consisrate_B_gender <- new_trfic_stps_B %>% group_by(Officer_Gender) %>% 
    summarise(rate = mean(consistent == 0))

tdf4 <- data.frame(Variable = rep("Gender", nrow(consisrate_B_gender)), 
                   Value = unlist(consisrate_B_gender[,2]), 
                          Profile = unlist(consisrate_B_gender[,1]))

# by race
lvls <- as.character(levels(new_trfic_stps_B$Officer_Race)[2:8])
consisrate_B_race <- new_trfic_stps_B %>% filter(Officer_Race %in% lvls) %>%
    group_by(Officer_Race) %>% 
    summarise(rate = mean(consistent == 0))


tdf5 <- data.frame(Variable = rep("Race", nrow(consisrate_B_race)), 
                   Value = unlist(consisrate_B_race[,2]), 
                          Profile = unlist(consisrate_B_race[,1]))

# by years of service
new_trfic_stps_B <- cbind(new_trfic_stps_B, Service_group = 
                        cut(new_trfic_stps_B$Officer_Years_of_Service, 
                            breaks =  c(0, 3, 10, 100), 
                            labels = c("Newbie", "Mid carreer", "Experienced")))

consisrate_B_yos <- new_trfic_stps_B %>% group_by(Service_group) %>% 
    summarise(rate = mean(consistent == 0))

tdf6 <- data.frame(Variable = rep("Service_group", nrow(consisrate_B_yos)), 
                   Value = unlist(consisrate_B_yos[,2]), 
                          Profile = unlist(consisrate_B_yos[,1]) )

# officer profiles
offcr_prof <- rbind(tdf4, tdf5, tdf6)
```

On overall consistency rate shown by CMPD is towards race of black is 
`r round(consisrate_B,2)`.

```{r, cache=T}
# plot by officer profile
offcr_prof$Profile <- as.character(offcr_prof$Profile)

ggplot(offcr_prof, aes(x = Variable, y = Value, group = 1, col = Profile)) + 
    geom_point() + geom_hline(aes(yintercept = consisrate, 
                                  col = "overall consistency rate"), 
                              size=0.25, type = "b") + 
    geom_hline(aes(yintercept = consisrate_B, 
                                  col = "overall consistency rate (Black)"), 
                              size=0.25, type = "b")

```

Graph above indicate that female police officers tend to be less consistent in their
action towards race of black. It can be seen that police officer with races from Native Hawaiian/Oth Pac Island, and American Indian/Alaska Native show more consistency towards race of black, whereas Hispanic/Latino police officers show less consistent approach. Also it can be seen that police officers who are new, shows more consistency towards race of black, where as experienced police officers have a less consistent approach. As pointed out previously it will interesting to look at where these inconsistancies are leaning towards, i.e. whether police offciers being strict or lenient towrds black groups.



